\textbf{zklora: Verifiable Multi-Party Inference for Evaluating Private LoRA Weights}

### 2025-01-12 paper/main.tex
convert table to figures

@@

We benchmarked \texttt{zklora} by measuring the creation and verification times for our verifiable computations across multiple base models and private LoRA configurations. For each model, we computed a forward pass for a batch of 3 items with sequence length 5. 
Each experiment simulates the workflow where the \emph{Base Model User} uses local data to evaluate the base model + LoRA combination, requesting remote LoRA computations from the \emph{LoRA Owner}. 
We report total and average proof-generation times (\texttt{total\_prove}, \texttt{avg\_prove}) and verification times (\texttt{total\_verify}, \texttt{avg\_verify}), as well as relevant model parameters.

\begin{table}[ht]
    \centering
    \small
    \begin{tabular}{l c c c c c c c}
    \toprule
    \textbf{base\_model} & \textbf{\#lora} & \textbf{total\_params} & \textbf{avg\_params} & \textbf{total\_prove} & \textbf{avg\_prove} & \textbf{total\_verify} & \textbf{avg\_verify} \\
    \midrule
distilgpt2 & 24 & 589824 & 24576.0 & 759.33 & 31.64 & 16.56 & 0.69 \\
gpt2 & 48 & 2359296 & 49152.0 & 1675.58 & 34.91 & 32.79 & 0.68 \\
Llama-3.2-1B & 32 & 851968 & 26624.0 & 991.93 & 31.00 & 24.91 & 0.78 \\
Llama-3.3-70B-Instr. & 80 & 11796480 & 147456.0 & 3749.76 & 46.87 & 123.11 & 1.54 \\
Llama-3.1-8B-Instr. & 32 & 5242880 & 163840.0 & 1527.40 & 47.73 & 35.79 & 1.12 \\
Mixtral-8x7B-Instr. & 32 & 10485760 & 327680.0 & 2357.61 & 73.68 & 44.30 & 1.38 \\
    \bottomrule
    \end{tabular}
    \caption{Benchmark results for \texttt{zklora} showing multi-party inference overhead on various model + LoRA combinations. Times are in seconds.}
    \label{tab:results}
\end{table}

Notably, although proof-generation times can grow with model size, verification remains comparatively fast (e.g., \texttt{avg\_verify} of about 1--2 seconds even for 70B-scale models). This result demonstrates that \texttt{zklora} allows frequent correctness checks without excessive overhead, which is crucial for real-world “trial runs.” In larger-scale usage scenarios, once trust in the LoRA module is established, \texttt{zklora} may no longer be needed.

%------------------------------------------
% 2. MULTI-PARTY INFERENCE METHOD
%------------------------------------------
\section{Multi-Party Inference: zklora for Verifiable Evaluation}
Our aim is to enable the \emph{Base Model User} to run a \emph{validation experiment} on their local data, incorporating the private LoRA module from the \emph{LoRA Owner}. The Base Model User can measure task performance (e.g., cross-entropy loss, accuracy) to decide whether the LoRA meets performance requirements.

Figure~\ref{fig:toy-mlp-lora} illustrates a simple scenario:
\begin{enumerate}
    \item \textbf{Base Model Forward Pass (Base Model User).}
    The \emph{Base Model User} runs layers without LoRA parameters locally, generating partial activations.

    \item \textbf{Forward to LoRA Adapter (Base Model User).}
    Whenever the next layer includes a LoRA transform, the \emph{Base Model User} sends these partial activations to the \emph{LoRA Owner}, who holds the private LoRA module.

    \item \textbf{LoRA Computation \& Proof Generation (LoRA Owner).}
    The \emph{LoRA Owner} applies the LoRA transformation to the activations and returns the updated activations alongside a \emph{cryptographic proof} of correctness. In our setup, \texttt{zklora} interfaces with \texttt{ezkl} to produce these proofs.

    \item \textbf{Resuming the Forward Pass (Base Model User).}
    The \emph{Base Model User} injects the updated activations back into the local model state and continues the forward pass.

    \item \textbf{Verification \& Validation (Base Model User).}
    The \emph{Base Model User} verifies the LoRA proof. If valid, the outputs from the combined base + LoRA model are guaranteed to match the LoRA Owner’s claimed computation. The Base Model User can now measure final metrics (loss, accuracy, etc.) to assess \emph{quality} on the target task.
\end{enumerate}

@@

\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{A simplified 2-layer MLP with LoRA on the first layer. The Base Model User (blue box) runs most computations locally but calls the LoRA Owner (dashed box) to update partial activations. After verifying correctness, the Base Model User can measure final metrics to decide whether to adopt the private LoRA more extensively.}
\label{fig:toy-mlp-lora}
\end{figure}

%------------------------------------------
% 3. INTRODUCTION
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) excel at a variety of tasks but come with significant computational and storage costs \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by introducing a small number of trainable parameters into the model, drastically reducing memory overhead for fine-tuning \cite{ding2022delta}.

When two entities, a \emph{LoRA Owner} and a \emph{Base Model User}, collaborate on a shared LLM, they may have conflicting constraints. The \emph{LoRA Owner} wants to keep custom LoRA weights private (e.g., for competitive or regulatory reasons), while the \emph{Base Model User} wants to \emph{test} or \emph{validate} these LoRA weights on a target task before deciding whether to deploy them at scale. Under normal circumstances, either the LoRA must be fully shared (sacrificing privacy), or the Base Model User must trust the LoRA Owner’s claims about performance with little evidence.

In this paper, we present \texttt{zklora}, a library that provides a \emph{verifiable computation} solution for the \emph{Base Model User} to \emph{evaluate} a private LoRA module. By orchestrating a multi-party inference, \texttt{zklora} allows the Base Model User to run the base model locally while sending partial activations to the LoRA Owner. Crucially, the LoRA Owner returns not only updated activations but also a cryptographic \emph{proof} that the correct LoRA transformation was applied. In this way, the Base Model User can confirm the final outputs truly reflect the LoRA module’s capabilities. After verifying and measuring results on a validation dataset, the Base Model User can later choose how to adopt the LoRA at larger scale.

@@

\subsection{Low-Rank Adaptation}
LoRA \cite{hu2021lora} has emerged as a popular way to efficiently fine-tune LLMs by injecting small, low-rank matrices into specific layers. This approach scales well, requiring less GPU memory than full fine-tuning. Various frameworks build on LoRA for GPT-2, GPT-Neo, LLaMA, and others \cite{ding2022delta}. By confining most updated parameters to low-rank adapter matrices, LoRA drastically reduces memory overhead during fine-tuning. As a result, domain-specific LoRA modules can be swapped in or out of the same base model, avoiding the need to retrain or store complete copies of large networks.

\subsection{Incrementally Verifiable Computation}
Valiant~\cite{valiant2008incrementally} introduced the notion of Incrementally Verifiable Computation (IVC), which offers updatable, compact proofs of correctness over multi-step computations. Nova~\cite{kothapalli2022nova} and HyperNova~\cite{kothapalli2024hypernova} refine these ideas, leveraging folding schemes and zero-knowledge properties to enable scaling of verifiable computations in multi-step neural network inference. IVC-based approaches align well with repeated computations needing multiple checkpoints of trust, such as multiple forward passes or partial training loops.

%------------------------------------------
% 5. CONCLUSION
%------------------------------------------
\section{Conclusion}
\texttt{zklora} enables a \emph{Base Model User} to verify the quality of a private LoRA module without revealing the LoRA weights. By pairing multi-party inference with cryptographic proofs, \texttt{zklora} ensures that the user can trust the computed outputs during validation experiments. After verifying performance, the Base Model User may adopt any approach they choose to integrate or deploy the private LoRA more widely, since the module has been proven effective. Our experiments show this approach is feasible across a range of model sizes, with \emph{fast verification times} minimizing the overhead of a verifiable evaluation phase. Future directions include further circuit optimization and exploring multi-owner scenarios, in which multiple private LoRA modules could be composed and verified within the same pipeline.

>>>>

%------------------------------------------
% Move INTRO to the beginning
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) excel at many tasks but can be expensive to store and fine-tune \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by injecting small adapter matrices, making specialized fine-tuning more efficient \cite{ding2022delta}.

However, if a \emph{LoRA Owner} wants to keep these weights private while a \emph{Base Model User} needs to validate them on a local dataset, a direct sharing of the LoRA parameters is impossible. \texttt{zklora} solves this dilemma by orchestrating multi-party inference: The Base Model User runs the base model layers locally, forwarding partial activations to the LoRA Owner for the private transform. A zero-knowledge proof is generated to demonstrate correctness, ensuring that final performance metrics are truly representative of the private LoRA.

@@

We benchmarked \texttt{zklora} by measuring the creation and verification times across multiple base models and private LoRA configurations. 
Figures~\ref{fig:settings}, \ref{fig:proof}, and \ref{fig:verify} show the total LoRA parameter count (horizontal axis) versus:
\begin{itemize}
    \item \textbf{Total Settings Time} (sec),
    \item \textbf{Total Proof Generation Time} (sec),
    \item \textbf{Total Verification Time} (sec),
\end{itemize}
respectively.

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.55\linewidth]{figs/fig_settings.pdf}
    \caption{Total LoRA params (in millions) vs. total settings time. Each point is a different model.}
    \label{fig:settings}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.55\linewidth]{figs/fig_proof.pdf}
    \caption{Total LoRA params (in millions) vs. total proof generation time. Each point is a different model.}
    \label{fig:proof}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[width=0.55\linewidth]{figs/fig_verify.pdf}
    \caption{Total LoRA params (in millions) vs. total verification time. Each point is a different model.}
    \label{fig:verify}
\end{figure}

Notably, while proof-generation times can grow with model size, verification remains consistently fast (on the order of 1--2 seconds per proof for large models). This helps real-world “trial runs,” letting the Base Model User confirm private LoRA quality without incurring excessive overhead.

%------------------------------------------
% 2. MULTI-PARTY INFERENCE METHOD
%------------------------------------------
\section{Multi-Party Inference: zklora for Verifiable Evaluation}

Our approach allows the \emph{Base Model User} to incorporate a private LoRA module from the \emph{LoRA Owner}:

\begin{enumerate}
    \item \textbf{Base Model Forward Pass.} 
    The user runs all layers without LoRA locally, producing partial activations.

    \item \textbf{Forward to LoRA Adapter.} 
    The partial activations are sent to the LoRA Owner for private transformations.

    \item \textbf{LoRA Computation \& Proof Generation.} 
    The LoRA Owner updates the activations and produces a zero-knowledge proof of correctness.

    \item \textbf{Resuming the Forward Pass.} 
    The updated activations are returned and merged into the local forward pass.

    \item \textbf{Verification \& Validation.} 
    The Base Model User verifies the proof and measures final metrics (e.g., accuracy) to decide if the LoRA meets their needs.
\end{enumerate}

@@

\node[draw, rounded corners, thick, color=blue!40!black,
      fit=(input)(W1)(act)(W2)(output),
      label={[font=\small, color=blue!40!black]above:\textbf{Base Model User}},
      inner sep=0.4cm
] (baseModelFit) {};
\end{tikzpicture}
\caption{A simplified 2-layer MLP with LoRA on the first layer. The Base Model User (blue box) runs most computations locally, then requests the LoRA Owner (dashed box) to update partial activations.}
\label{fig:toy-mlp-lora}
\end{figure}

%------------------------------------------
% 3. HOW WE GENERATE THE ZERO-KNOWLEDGE PROOF
%------------------------------------------
\section{Generating the Zero-Knowledge Proof}

Central to \texttt{zklora} is the cryptographic proof that the LoRA computations were conducted faithfully. Below is a high-level outline:

\paragraph{Circuit Compilation (Model + JSON Input).}
The LoRA Owner provides an ONNX model that includes the LoRA parameters. We compile it into a “circuit” representing the layer computations. Next, the Base Model User's partial activations, saved as JSON, feed into this circuit as inputs.

\paragraph{Settings and Key Generation.}
We produce (1) a “settings” file that includes circuit metadata, (2) a proving key (PK) and verification key (VK), and optionally (3) a structured reference string (SRS) for polynomial commitments.

\paragraph{Witness Creation.}
We run the partial activations through the circuit, storing all intermediate wire values in a “witness” file that reflects the correct usage of the LoRA layer.

\paragraph{Proving.}
Using the PK, the witness, and the circuit, the LoRA Owner generates a proof file. This file certifies that the LoRA transformation was done correctly, matching the declared ONNX logic.

\paragraph{Verification.}
The Base Model User loads the proof and verifies it with the VK (plus SRS). If it checks out, the user trusts that the LoRA updates are accurate, without ever seeing the LoRA weights directly.

@@

\subsection{Low-Rank Adaptation}
LoRA \cite{hu2021lora} is widely adopted for parameter-efficient fine-tuning across GPT-2, GPT-Neo, and LLaMA. By focusing updates on low-rank adapters, it reduces memory overhead \cite{ding2022delta}.

\subsection{Incrementally Verifiable Computation}
Earlier cryptographic frameworks \cite{valiant2008incrementally, kothapalli2022nova, kothapalli2024hypernova} show how to prove correctness of multi-step computations. \texttt{zklora} combines these proof techniques with a multi-party inference flow, ensuring that LoRA transformations remain private yet verifiable.

%------------------------------------------
% 5. CONCLUSION
%------------------------------------------
\section{Conclusion}

We present \texttt{zklora} as a verifiable multi-party inference solution to validate private LoRA adapters. Our benchmarks show that proof generation scales with model size, while verification remains consistently quick. This enables the Base Model User to run iterative experiments on LoRA modules without risking the LoRA Owner’s privacy. Future extensions might explore multi-owner LoRA composition or advanced multi-party computation frameworks that guard both user data and model parameters with cryptographic guarantees.
